---
title: "R Basics Part 2"
author: "INSOFE"
date: "2 June 2018"
output: 
  html_document:
    toc : true
    toc_float : true
    toc_depth : 5
---
# Basic Preprocessing steps in R
## Agenda 
* Setting the Working Directory and knowing the Current working Directory
* Reading data from a csv file and Creating a Data Frame 
* Head and Tail functions
* Getting the Column names and row names
* Renaming the Columns
* Checking the data types of columns - is.* and str functions
* Converting the data into required data types - as.* function
* Subsetting the Data Frame 
* Checking for the missing values
* Imputing the missing values - Central Imputation
* Creating a new column -- Power to Weight ratio
* Numerical to Categorical - Binning
* Categorical  to Numerical - Dummifying
* Scaling the Numerical Attributes
* Creating Functions
* Using the apply family 
* Reading data from two files and merging 

## Present Working Directory -- getwd()
```{r}
# getwd()
```
### Setting the working directory -- setwd()
```{r}
# setwd()
```

## Reading data from a csv file and Creating a Data Frame
   - Dataset mtcars is shared in the form of a .csv fil, we will read this in to R evnironment
```{r}
mtcars = read.csv(file = "mtcars.csv",header=TRUE,sep=",")

class(mtcars)
```

#### Data set information 
  
  The mtcars dataset was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 
  10 aspects of automobile design and performance for 32 automobiles (1973–74 models).."The data concerns 
  city-cycle fuel consumption in miles per gallon."

#### Description of the Attributes 

-	mpg	Miles/(US) gallon
-	cyl	Number of cylinders
-	disp	Displacement (cu.in.)
-	hp	Gross horsepower
-	drat	Rear axle ratio
-	wt	Weight (1000 lbs)
-	qsec	1/4 mile time
-	vs	Engine (V-shaped,straight)
-	am	Transmission (0 = automatic, 1 = manual)
-	gear	Number of forward gears
-	carb	Number of carburetors


## Head and Tail functions

### Let's take a look at top rows of the dataframe
```{r}
head(mtcars)
```

### Let's take a look at the bottom rows  of the dataframe
```{r}
tail(mtcars)
```

## Getting the Column names and row names

### Column names
```{r}
colnames(mtcars)
```
### Row names
```{r}
rownames(mtcars)
```

## Renaming the  Rows and Columns
### Renaming the rows
```{r}
rownames(mtcars) = mtcars$X
rownames(mtcars)
mtcars$X=NULL
mtcars
```

###  Renaming the  Columns
```{r}
colnames(mtcars)[6] = "weight"
```

## Checking the data types of columns - is.* and str functions
### Data type of the Column 
```{r}
# Data type of the Column checked by is.* will return in a logical vector of length 1
is.numeric(mtcars$Cars)
is.character(mtcars$hp)
```

### Data types of the Columns in the entire data set
```{r}
str(mtcars)
```
## Converting the data into required data types - as.* function
```{r}

mtcars$cyl = as.factor(mtcars$cyl)
# converting am column to factor 
mtcars$am = as.factor(mtcars$am)

```

##  Subsetting the Data Frame
### Subsetting based on column names
#### Subsetting based on a single column by name
```{r}
mtcars[,'carb']

```
#### Subsetting based on a single column by column number (also referred to as column index)
```{r}
mtcars[,11]

```
#### Subsetting based on a two or more columns based on column name
- Observe the class of the object returned in both the cases 
```{r}
mtcars[,c('mpg','carb')]
```
#### Subsetting based on a two or more columns based on column index number

```{r}

mtcars[,c(1,11)]
class(mtcars[,c(1,11)])

```

#### Subsetting based on index row name
```{r}
rownames(mtcars)
mtcars['Duster 360',]
class(mtcars['Duster 360',])
```
#### Subsetting based on index row number
```{r}
mtcars[7,]
class(mtcars[7,])
```

#### Subsetting based on multiple index values 
```{r}
mtcars[c(6,7),]
class(mtcars[c(6,7),])
```
#### Subsetting using the subset function
```{r}
#Subsetting data
s1 <-subset(mtcars,select=c(cyl,mpg))
s1
```

#### Subsetting based on multiple index values and column names
```{r}
mtcars[c('Duster 360','Valiant'),c('mpg','hp')]
class(mtcars[c('Duster 360','Valiant'),c('mpg','hp')])
```


#### Subsetting based on multiple index values and column names

```{r}
mtcars[c(6,7),c(1,2)]
class(mtcars[c(6,7),c(1,2)])
```

## Checking for the missing values
### Checking by using the for loop and *is.na* fuction
```{r}
colsum=c()
for (i in colnames(mtcars))
{
  sum = 0
  for (j in mtcars[,i])
  {
    if(is.na(j)==TRUE)
    {
      sum = sum+1
    }else
    {
      sum = sum
    }
  }
  colsum[i] = sum
}

colsum
```

### Checking by using the colSums
```{r}
colSums(is.na(mtcars))
```

### Getting the Sum of the total NA values in the Dataset
```{r}
sum(is.na(mtcars))
```

## Imputing the missing values - Central Imputation
```{r}
library(DMwR)
mtcars_imputed  = centralImputation(mtcars)
colSums(is.na(mtcars_imputed))
```

## Creating a new feature power to weight ratio
- creating a new feature and adding it to the existing dataframe

```{r}
mtcars_imputed$pwr <-mtcars_imputed$hp/mtcars_imputed$weight 
head(mtcars_imputed)
```


## Numerical to Categorical - Binning
### Manual binning
```{r}
mtcars_imputed $hp
# After looking the numbers we can get the idea of how to bin them 
bins = cut(mtcars_imputed $hp,breaks = seq(0,500,20),include.lowest = T,right = T)
bins
# Count the number of rows that comes under a particular bin
table(bins)

```

### Use the *discretize* function from infotheo package
```{r}
library(infotheo)
(nbins = floor(sqrt(NROW(mtcars_imputed $hp))))
discretize(mtcars_imputed $hp,disc = "equalwidth",nbins)
```

```{r}
discretize(mtcars_imputed $hp,disc = "equalfreq",nbins)
```

```{r}
str(mtcars_imputed)
```

## Categorical to Numerical 

```{r}
library(dummies)
dummy(mtcars_imputed$vs)
```

## Scaling the Numerical Attribute
```{r}
library(vegan)
mtcars_imputed_standardized <- decostand(x =mtcars_imputed[,c('mpg','disp','drat','weight','qsec','gear','carb')],method ="range",MARGIN = 2)
str(mtcars_imputed_standardized)
range(mtcars_imputed_standardized$mpg)
```

## Functions in R
  - Function syntax
   funcName <- function (argument) {
     statements
   }
   
#### Define and Call a function 
- paste function is used for string concatenation and string formatting for the statements we print on console
```{r}
pow <- function(x, y = 3) #Default Values for y Arguments
  {
  result <- x^y
  print(paste(x, "raised to the power", y, "is", result))
}
pow(3)
```


```{r}
pow(3) # takes the default y value

pow(8, 2) # we can pass the arguments , r considers them in the same order and passes them in the place of prameters

pow(x = 8, y = 2) # Named Arguments 

pow(y = 2, x = 8)
```

#### R Return Value from Function
```{r}
check = function(x) {
  if (x > 0) {
    result <- "Positive"
  }
  else if (x < 0) {
    result <- "Negative"
  }
  else {
    result <- "Zero"
  }
  return(result)
}

check(1)
```


#### Functions without return()
```{r}
# If there are no explicit returns from a function, the value of the last evaluated expression is returned automatically in R.

check <- function(x) {
  if (x > 0) {
    result <- "Positive"
  }
  else if (x < 0) {
    result <- "Negative"
  }
  else {
    result <- "Zero"
  }
  result
}

check(-1)

```


#### Multiple Returns
- The return() function can return only a single object. If we want to return multiple values in R, we can use a list (or other objects) and return it.

```{r}
multiReturn = function() {
  myList <- list(1, 20, c("a","b"))
  return(myList) 
}

multiReturn()
```

## Family of *apply* functions 
- We can do some very complex operations with using the se handy functions from the *apply* family of functions
- apply functions take a list/vector or dataframe as input, apply a function on the input and produce the output, let us look a some of the apply functions and how to use them.

#### *lapply*
- lapply takes a list as an input, and returns a list as the output.

- syntax :  lapply(X,FUN)



#### mean function on a column in the dataframe  mtcars
```{r}
lapply(mtcars['mpg'],mean)
class(lapply(mtcars['mpg'],mean))
```


#### *sapply*
- sapply takes a list as an input, and returns a vector as output.

- syntax :  sapply(X,FUN)

#### min and max function on a column in the dataframe  mtcars
```{r}
sapply(mtcars['mpg'],max)
class(sapply(mtcars['mpg'],max))
```

#### *tapply*
- tapply splits the array based on specified data, usually factor levels and then applies the function to it.

- syntax :  tapply(X,Y,FUN)
- tapply splits X based on levels in Y


#### mean function on a column in the dataframe  mtcars
```{r}
str(mtcars)
```

```{r}
tapply(mtcars$weight, mtcars$cyl, mean)
```

#### Find the max hp of cars based on number of cylinder
```{r}
tapply(mtcars$hp,mtcars$cyl,max)
```

#### *apply*
- apply takes input
  - 1. Dataframe
  - 2. An argument which specfies if we want to apply the funtions row wise or column wise and
  - 3. The function we want to apply

- syntax :  apply(X,1 (or 2) ,FUN)
  - 1 for rows
  - 2 for columns



#### apply the max function on rows
```{r}
colnames(mtcars_imputed)
colnames(mtcars_imputed[,-c(1)])
mtcars_imputed[,-c(1)]
apply(mtcars_imputed[,-c(1)],1,max)
```
#### apply the mean  and max function on columns
```{r}
apply(mtcars_imputed[,-c(1)],2,max)
apply(mtcars_imputed[,-c(1)],2,max)
class(apply(mtcars_imputed[,-c(1)],2,max))
```



## Reading Data from two files and merging the data

```{r}
cust_bank_details = read.table("Customer_Bank Details_MV.csv",sep = ",",header = T)
cust_demographics = read.table("Customer_Demographics_MV_DOB.csv",sep = ",",header = T)
```

```{r}
head(cust_bank_details)
```

```{r}
colnames(cust_demographics)[1] = "ID"
head(cust_demographics)
```

#### Combining/Merging/Joining data from multiple files.

```{r}
library(data.table)
merged_df = merge(cust_bank_details,cust_demographics,by.x="ID",by.y = "ID")
head(merged_df)
```
