---
title: "Simple Linear Regression"
author: "INSOFE Lab"
date: "8th July, 2018"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

## Simple Linear Regression Model

* Clear environment variables

```{r}
rm(list=ls(all=TRUE))

```

## Read or load data
* Set working Directory and read the data from csv file

```{r}

#setwd("/home/seemabasantani/Seema Basantani/Academics/Batch44/CSE 7302c/Batch 44 - CSE 7302c/Simple Linear Regression")
setwd("D:/Batch 44/CSE 7302c/Class")
cars_data = read.csv(file="cars.csv", header=T)   # Read the data from cars.csv file

```

* Summary of the data and look for any missing values:

```{r}

names(cars_data)      # Display the column names
str(cars_data)        # Structure of the dataset
summary(cars_data)    # Summary of the dataset

```

* No missing values

## Scatter Plot
```{r}
plot(cars_data$speed,cars_data$dist,xlab="Speed in miles per hour",ylab="Distance in feet",main="Stopping Distance Vs. Speed", col= "blue") # Plot the dependent and independent variables

```

## Covariance between the attributes
```{r}

cov(cars_data)      # Covariance between independent and dependent variable

```
* The covariance of the speed of the car and distance is 109.9469

## Correlation between the attributes

```{r}
cor(cars_data$speed, cars_data$dist)      # Correlation between independent and dependent variable

```


* The correlation coefficient of the Speed of car and Distance is 0.8068949. 
* Since the value is close to 1 and has a +ve sign, we can conclude that the variables are positively correlated.

## Building the Linear Regression Model
* lm function is used to fit linear models
 
```{r}

LinearReg = lm(dist ~ speed, data = cars_data)
coefficients(LinearReg)

```

### Read the model summary

```{r}
## Summary of the linear model:
summary(LinearReg)

```


### Line of Best Fit

```{r}
plot(cars_data$speed,cars_data$dist,xlab="Speed in miles per hour",ylab="Distance in feet",main="Stopping Distance Vs. Speed: Best fit line", col= "blue")
abline(LinearReg,col="steelblue",lty=1,lwd=4)       # The function adds straight line to a plot

```

## Extracting residuals and fitted values

```{r}

# To extract the residuals:
LinearReg$residuals

# To extract the predictions
LinearReg$fitted.values

```

## Validity of linear regression assumptions

```{r}

par(mfrow = c(2,2))
plot(LinearReg)           # Check for validity of linear regression assumptions
par(mfrow = c(1,1))       # Resetting the plot layout

```




