---
title: "Activity-june-9"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
```{r echo=FALSE}
rm(list=ls())
setwd('/Users/bsriram/datasci/june/wk2/stats')
```

## Activity -1
 	A company has 10 employees and their average salary is 1 lakh per annum. Now they have recruited a CEO who salary is normally very high. Which of the following measure of central tendencies will be impacted
   **a.	Median**
   b.	Mode
   c.	Mean

### Answer: **Median**

## Activity -2
	Here is the data of experience of individuals in a CPEE class. We have grouped individuals into 6 groups and here is the data. Compute the average and median values for each group and list your observations.

### Answer:
```{r echo=FALSE}

experience = read.table("group.csv",sep = ",",header = T, na.strings=c('NA','?'))
```

###Questions:
1.	What is the average in each group?
```{r echo=FALSE}
mean(experience)
```
2.	What is the median in each group?
```{r echo=FALSE}
```
3.	What is the average experience across all groups?
```{r echo=FALSE}
```
4.	Does the average of the each group average experience is same as the Grand average ?
```{r echo=FALSE}
```


## Activity -3

```{r echo=FALSE}
library(modes)

deliverySvcs = read.table("deliveryserivces.csv",sep = ",",header = T, na.strings=c('NA','?'))
#print(deliverySvcs)

stats = function(x)
{
   res = c("mean" = mean(x),
   "med" = median(x),
   "q1" = quantile(x)[2],
   "q2" = quantile(x)[3],
   "q3" = quantile(x)[4],
   "min" = range(x)[1],
   "max" = range(x)[2],
   "iqr" = IQR(x),
   "var" = var(x),
   "sd" = sd(x))
   return(res)
}

apply(deliverySvcs, 2,stats)
```

## 6.	Analyze the performance of a class 
```{r echo=FALSE}
 Scores<-c( 11, 7.5, 8.5, 10, 10, 10.5, 5.5, 10, 9, 9.5, 5.25, 8, 6.5, 10.5, 8.75, 0, 6, 6,  6.75,8.75, 0, 9.5, 7.5, 8.5, 7)

stats(Scores)
boxplot(Scores)

plot
```
a.	How is the spread of the scores? Compute range, variance & standard deviation
b.	Find the 25th percentile, 50th percentile and 75 percentiles for this data
c.	Find the Q1, Q2 and Q3 and IQR for this data.
d.	Find outliers, if any. Do a boxplot to visualize the same.

```{r}
Supplier1 <-c(17,22,22,22,27)
Supplier2<-c(17,19,20, 27,27)

stats(Supplier1)
stats(Supplier2)
```
