---
title: "BostonHousingData_MultipleLinearRegression"
author: "INSOFE Lab"
date: "14th July 2018"
output:
  html_document:
    toc : yes
    toc_depth : 5
    toc_float : yes
---
# Agenda
 - Build a Multiple Linear Regression model
 - Analysis of the model
```{r}
# First things first , clean your R environment
rm(list = ls(all=TRUE))
# setwd("")       # Set the working directory
```
# Loading Libraries

```{r}
library(corrplot)
library(DMwR)
```

# _*Multiple Linear Regression Model*_
# Read or load data
```{r}
dataFile = "boston.csv"
boston_housing_data = read.csv(dataFile)
attach(boston_housing_data) # by attaching the dataframe to the memory, we can directly use the variable names

```
# Data Desctiption 

Data description
The Boston data frame has 506 rows and 14 columns.

This data frame contains the following columns:

- CRIM : per capita crime rate by town.

- ZN : proportion of residential land zoned for lots over 25,000 sq.ft.

- INDUS : proportion of non-retail business acres per town.

- CHAS : Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).

- NOX : nitrogen oxides concentration (parts per 10 million).

- RM : average number of rooms per dwelling.

- AGE : proportion of owner-occupied units built prior to 1940.

- DIS : weighted mean of distances to five Boston employment centres.

- RAD : index of accessibility to radial highways.

- TAX : full-value property-tax rate per $10,000.

- PT : pupil-teacher ratio by town.

- B : 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.

- LSTAT : lower status of the population (percent).

- MV : median value of owner-occupied homes in $1000s.

# EDA - Exploratory Data Analysis and Preprocessing steps
1. Look at the number of rows and columns
2. Column names
3. Look at the structure of the dataset using the function `str()`
4. Look at the summary of the dataset using the `summary()` 
5. Checking for 'NA' values

## 1. Look at the number of rows and columns

```{r}
dim(boston_housing_data)
```

## 2. Column names

```{r}
colnames(boston_housing_data)
```

## 3. Look at the structure of the dataset using the function `str()`

```{r}
str(boston_housing_data)
```

## 4. Look at the summary of the dataset using the `summary()` 

```{r}
summary(boston_housing_data)
```

## 5. Checking for 'NA' values
```{r}
sum(is.na(boston_housing_data))
```
## Converting a varible explicity to factor

```{r}
boston_housing_data$CHAS_F = factor(boston_housing_data$CHAS)
#dummies::dummy(boston_housing_data$CHAS_F)
boston_housing_data$CHAS_F1 = dummies::dummy(boston_housing_data$CHAS_F)[,1]
boston_housing_data$CHAS_F = NULL

```

## Scatter plot with a few selected attributes

```{r}
pairs(~CRIM+ZN+INDUS+NOX+RM+LSTAT,data=boston_housing_data,main="Scatterplot matrix with selected attributes")
```

## Scatter plot with all attributes

```{r}
pairs(~.,data=boston_housing_data,main="Scatterplot matrix")
```

## Correlation

```{r}
correlation_XPairwise = cor(boston_housing_data[, !(colnames(boston_housing_data) %in% c("MV","CHAS_F1"))])
corrplot(correlation_XPairwise)
```


### Write the pairwise correlation matrix to a csv file for later analysis

```{r}
write.table(correlation_XPairwise,file="pairwiseCorrelations1.csv",row.names=FALSE,col.names=FALSE,sep=",")
```

# Model building

## Simple Linear Regression 

```{r}
lineFitSinglePredictor = lm(MV ~ LSTAT, data = boston_housing_data); # fit a simple linear regression model
summary(lineFitSinglePredictor)
```

### Plot the scatter plot and the fitted line for single variable

```{r}
plotTitle = paste("Boston housing dataset : \n Median House Value vs %Lower Status Population\n Correlation coefficient = ",toString(cor(boston_housing_data$MV,boston_housing_data$LSTAT)),"\n MV = ",toString(lineFitSinglePredictor$coefficients[1]), "+ LSTAT*(",toString(lineFitSinglePredictor$coefficients[2]),")\n\n",sep=" ")
plot(boston_housing_data$LSTAT,boston_housing_data$MV,main=plotTitle,xlab="%Lower Status Population (LSTAT)",ylab="Median House Value (MV)")
abline(lineFitSinglePredictor,col="steelblue",lty=1,lwd=4)
```

### Summary of the model

```{r}
summary(lineFitSinglePredictor)
```

### Residual plots for the models

```{r}
plot(lineFitSinglePredictor,main="Linear model with a single predictor") 
```

### Cook's Distance plot for simple linear regression

```{r}
cd = cooks.distance(lineFitSinglePredictor)
plot(cd,main=paste("boston_housing_data"," : \n Cook's distances for each sample for the Simple Linear Model",sep=" "),xlab="Index",ylab="Cook's distance")
grid(10,10,lwd=2)

```

##  Multiple Linear Regression

### Build Model

```{r}

lineFitMultiplePredictors = lm(boston_housing_data$MV ~ B+CRIM+RM+LSTAT)
```

### Summary of the model

```{r}
summary(lineFitMultiplePredictors) 
```

### Residual plots for the multiple linear regression model

```{r}
plot(lineFitMultiplePredictors,main="Linear model with a multiple predictor") 

```

### Regression using all predictors

```{r}
lineFit_All = lm(MV ~ ., data = boston_housing_data) 
summary(lineFit_All)
```

### Regression using interaction terms
```{r}
lineFit_Interaction = lm(MV ~ LSTAT * AGE, data = boston_housing_data) 
summary(lineFit_Interaction)
plot(lineFit_Interaction)
```

### Regression using non-linear transformation of predictors

```{r}
lineFit_NonLinearTransform = lm(MV ~ CRIM + NOX + RM + DIS + PT + B + LSTAT + I(LSTAT ^ 2), data=boston_housing_data) 

```

### Regression using non-linear transformation of predictors along with a categorical predictor

```{r}
lineFit_NonLinearTransform_Categorical = lm(MV ~ CRIM + CHAS_F1 + NOX + RM + DIS + PT + B + LSTAT + I(LSTAT ^ 2), data=boston_housing_data) 
summary(lineFit_NonLinearTransform_Categorical)
plot(lineFit_NonLinearTransform,main="Linear model with multiple predictors and non-linear transformations") 

```

### Predictions on test data

```{r}
set.seed(123) # for reproducible results
sample.size <- floor(0.75 * nrow(boston_housing_data))
train.index <- sample(seq_len(nrow(boston_housing_data)), size = sample.size)
train <- boston_housing_data[train.index,]
test <- boston_housing_data[- train.index,]
testPredictions_simple_linear_model = predict(lineFitSinglePredictor, test, interval = "confidence") 
testPredictions_multiple_linear_model = predict(lineFit_All, test, interval = "confidence") 
```

## Evaluation of models

### Simple Linear Regression

```{r}
regr.eval(trues = test[,c("MV","LSTAT")],preds = testPredictions_simple_linear_model)
```


### Mutiple Linear Regression

```{r}
regr.eval(trues = test,preds = testPredictions_multiple_linear_model)

```









































