---
title: "Association rules"
author: "INSOFE Lab"
date: "28 July, 2018"
output: html_document
---

## Data preprocessing 

Clearing environment variables
```{r}
rm(list=ls(all=TRUE))

``` 

Loading required libraries

```{r}
# install.packages('arules')
library(arules)

```

Data loading

```{r}
data("Groceries")
Groceries
head(Groceries)

# Groceries <- read.transactions(file.choose())     # Read transactions from file

# Writing Groceries Transaction data to csv file
# Groceries_new <- as(Groceries,"data.frame")
# write.csv(Groceries_new, file = "Groceries.csv",row.names = TRUE)

```

Viewing the data

```{r}
# Groceries@itemInfo
inspect(Groceries[1:10])

```

Summarizing the data in transaction table

``` {r}
summary(Groceries)
#df=data.frame(Groceries)
#head(df)
```

Looking at the item frequency

``` {r}
itemFrequency(x = Groceries, type = "absolute")

```

```{r}
itemFrequency(x = Groceries, type = "relative")

```

Plotting the item frequency for top 15 items

```{r, fig.width=6}
itemFrequencyPlot(x = Groceries, topN = 20, type="relative")

```

## Mining rules/pattern
You will always have to pass the minimum required support/confidence or both

We set the minimum support to 0.001 and minimum confidence of 0.8

Get the rules :

```{r}
rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8), control = list(verbose = F))
length(rules)

```

Inspecting top 5 rules by rounding off all the numerical values to 2 digits in the output

```{r}
getOption("digits")
options(digits = 2)
inspect(rules[1:5])
```

Reading the pattern is now easier.
for e.g. : If someone buys yogurt and cereals, they are 81% likely to buy Whole milk too.

We can get summary information about the rules that give us some interesting information such as:
* The number of rules generated: 410
* The distribution of rules by length: Most rules are 4 items long
* The summary of quality measures: interesting to see ranges of support, lift, and confidence.
* The information on the data mined: total data mined, and minimum parameters.

```{r}
summary(rules)

```

when we have large set of rules then we would prefer to sort the rules based on requirement and take decision using top rules

Let's say we wanted to have the most likely rules, we can easily sort out the rules by confidence in descending order by using following piece of code

```{r}
sorted_rules <- sort(rules, by = "confidence" ,decreasing = T)
inspect(sorted_rules[1:5])

```
## Visualization 
* Visualizing the rules using a graph

```{r, fig.width=16, fig.height=12}

sorted_rules <- sort(rules, by = "support" ,decreasing = T)
inspect(sorted_rules[1:5])

library(arulesViz)
plot(sorted_rules[1:5], method = "graph", shading = NA, interactive = F)

```

Perhaps "sometimes rules with length greater than 3 may not be interesting for client"
Let's say we want more concise rules with length not exceeding 3, which can be easily achieved by adding "maxlen" parameter to the apriori function.

```{r}
rules_maxlen3 <- apriori(data = Groceries, parameter = list(supp = 0.001, conf = 0.8, maxlen = 3), control = list(verbose = F)) 
inspect(rules_maxlen3)

```

Redundancies 

* Rules might be repeating with different combinations with similar support and confidence values e.g. {A,B} -> {C}, {A,C} -> {B}
* every subset of a frequent itemset is also frequent (Downward closure property)

We can check and remove the redundant rules by using below code snippet
```{r}
sum(is.redundant(sorted_rules))
sorted_unique_rules <- sorted_rules[!is.redundant(sorted_rules)]
sum(is.redundant(sorted_unique_rules))

sum(is.redundant(rules_maxlen3))
```

## Targetting items

Now that we know to generate rules and limit the output.

Let's say we want to target items and generate rules.

Two types of targets which the client might be interested in, let us explore those by using an example of "whole milk"
* What are customer likely to buy before buying "whole milk"
* What are customer likely to buy if they purchase "whole milk"

essentially you are asking will i be able to set my lhs, rhs or both in my rules?

well! Lets look at the first target condition
** What are customer likely to buy before buying "whole milk" **
```{r}
target_rules1 <- apriori(data = Groceries, parameter = list(supp = 0.001, conf = 0.8, minlen = 2), appearance = list(rhs = "whole milk"), control = list(verbose = F))

sorted_target_rules1 <- sort(target_rules1, by = "confidence", decreasing = T)
inspect(sorted_target_rules1[1:10])
```

** What are customer likely to buy if they purchase "whole milk" **
```{r}
target_rules2 <- apriori(data = Groceries, parameter = list(supp = 0.001, conf = 0.15, minlen = 2), appearance = list(lhs = "whole milk"), control = list(verbose = F))

sorted_target_rules2 <- sort(target_rules2, by = "confidence", decreasing = T)
inspect(sorted_target_rules2)
length(sorted_target_rules2)
```

## Visualization

* Visualizing the rules using a graph

```{r, fig.width=16, fig.height=12}

plot(sorted_target_rules2[1:6], method = "graph", shading = NA, interactive = F)

```